Transfer Learning in Visual and Relational Reasoning

### 2020-02-14 Transfer_Learning/arxiv_version/submission_folder/v2/experiments.tex
arxiv v2 submission files

\begin{figure}[b!]
	\centering
	\includegraphics[width=0.5\textwidth]{../img/plots/clevr_baselines.pdf}
	\caption{Comparison of SAMNet to baselines on CLEVR.}
	\label{fig:clevr_baselines}
\end{figure}

We implemented and trained our SAMNet model using MI-Prometheus~\cite{kornuta2018accelerating}, a framework based on PyTorch~\cite{paszke2017automatic}.
We evaluated the model on
the CLEVR dataset~\cite{johnson2017clevr}, a diagnostic dataset for Image Question Answering and on the COG dataset~\cite{yang2018dataset}, a video reasoning~\cite{mogadala2019trends} dataset developed for the purpose of research on relational and temporal reasoning.
We briefly describe both datasets in the following two subsections.
In all experiments, we trained SAMNet using $k = 8$ reasoning steps and external memory with $N = 8$ slots, each storing an array of 128 floats.
When working with CLEVR, we deactivated SAMNet's external memory and temporal-related modules as they are unnecessary while reasoning about static images. 
Supplementary information regarding the dataset parameters and experiments can be found in the appendix.

@@

\begin{figure*}[!t]
	\centering
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../img/plots/cog_canonical_baseline_no_labels.pdf}
	\end{subfigure}%
	\newline
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../img/plots/cog_hard_baseline_labels.pdf}
	\end{subfigure}%
	\caption{Comparison of test set accuracies of SAMNet (blue) with original results achieved by the baseline model~\cite{yang2018dataset} (dotted gray) on Canonical (top) and Hard (bottom) variants of the COG dataset.}
	\label{fig:samnet_cog_detailed}\vspace{-10pt}
\end{figure*}

@@

\cref{fig:clevr_baselines} presents the accuracy achieved by SAMNet when compared to selected SOTA models, namely: Transparency by Design networks (TbD)~\cite{mascharka2018transparency}, Feature-wise Linear Modulation (FiLM)~\cite{perez2018film} and the Program Generator + Execution Engine (PG + EE)~\cite{johnson2017inferring}.
SAMNet reaches a comparable 96.2\% accuracy, with TbD achieving the best score of 98.7\%.

@@

COG dataset contains short videos, with associated questions grouped into 23 categories~\footnote{COG also proposes a pointing version of the tasks, which are not considered in this work.}.
COG comes in two variants: Canonical (easy) and Hard, differing mainly on the number of frames in the video, the maximum amount of look-back in frame history containing relevant information for reasoning, and the number of object distractors present in a given frame (see~\cref{tab:cog_variants}).
Since the number of task classes is large, we designed a 2-level hierarchy of task groups using the
description of tasks, as shown in~\cref{fig:task-groups}.

@@

\begin{figure}[b!]
	\centering
	\includegraphics[width=0.7\textwidth]{../img/architecture/hierarchy}
	\caption{Hierarchy of Task Groups in the COG dataset.}
	\label{fig:task-groups}
\end{figure}

@@

We compared our results with the baseline model introduced in the same paper as the COG dataset~\cite{yang2018dataset}.
The most important results are highlighted in~\cref{fig:samnet_cog_detailed}; full comparison can be found in the supplementary material.

@@

In order to quantify SAMNet's feature transfer ability, we used the CoGenT variants of CLEVR and performed a set of experiments, starting from training on CoGenT-A, followed by:
an immediate test (zero-shot learning) on CoGenT-B; and fine-tuning for a single epoch on 30k samples from CoGenT-B before testing (following the methodology used in~\cite{johnson2017inferring, mascharka2018transparency, perez2018film, marois2018transfer}).
In \cref{fig:CoGenT-B-results}, we compare the achieved performance with the previously selected SOTA models.
We also consider a direct test on CoGenT-A; and a test on CoGenT-A after fine-tuning on B.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.9\textwidth]{../img/plots/cog_temporal_transfer_baselines.pdf}
	\caption{Temporal transfer capabilities of SAMNet (blue) and baseline models (gray -striped and -dotted) when moving from Canonical to Hard  in COG.}
	\label{fig:samnet_cog_overall_transfer}
\end{figure}

>>>>

We implemented and trained our SAMNet model using MI-Prometheus~\cite{kornuta2018accelerating}, a framework based on PyTorch~\cite{paszke2017automatic}.
We evaluated the model on
the CLEVR dataset~\cite{johnson2017clevr}, a diagnostic dataset for Image Question Answering and on the COG dataset~\cite{yang2018dataset}, a video reasoning~\cite{mogadala2019trends} dataset developed for the purpose of research on relational and temporal reasoning.
We briefly describe both datasets in the following two subsections.
In all experiments, we trained SAMNet using $k = 8$ reasoning steps and external memory with $N = 8$ slots, each storing an array of 128 floats.
When working with CLEVR, we deactivated SAMNet's external memory and temporal-related modules as they are unnecessary while reasoning about static images. 
Additional information regarding the dataset parameters 
can be found in~\cref{sec:datasets-desc}.

@@

@@

\Cref{fig:clevr_baselines} presents the accuracy achieved by SAMNet when compared to selected state-of-the-art models: Transparency by Design networks (TbD)~\cite{mascharka2018transparency}, Feature-wise Linear Modulation (FiLM)~\cite{perez2018film} and the Program Generator + Execution Engine (PG + EE)~\cite{johnson2017inferring}.
SAMNet reaches a comparable 96.2\% accuracy, with TbD achieving the best score of 98.7\%.

\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.5\textwidth]{../img/plots/clevr_baselines.pdf}
	\caption{Comparison of SAMNet to baselines on CLEVR.}
	\label{fig:clevr_baselines}
\end{figure}

@@

COG dataset contains short videos, with associated questions grouped into 23 categories\footnote{COG also proposes a pointing version of the tasks, which are not considered in this work.}.
COG comes in two variants: Canonical (easy) and Hard, differing mainly on the number of frames in the video, the maximum amount of look-back in frame history containing relevant information for reasoning, and the number of object distractors present in a given frame (see~\cref{tab:cog_variants}).

@@

Since the number of task classes is large, we designed a 2-level hierarchy of task groups using the
description of tasks, as shown in~\cref{fig:task-groups}.

\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.7\textwidth]{../img/architecture/hierarchy}
	\caption{Hierarchy of Task Groups in the COG dataset.}
	\label{fig:task-groups}
\end{figure}

@@

We compared our results with the baseline model introduced in~\cite{yang2018dataset}, the same paper as the COG dataset.
The most important results are highlighted in~\cref{fig:samnet_cog_detailed}; full comparison can be found in~\cref{sec:cog-all-results}.

\begin{figure}[htbp]
	\centering
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../img/plots/cog_canonical_baseline_no_labels.pdf}
	\end{subfigure}%
	\newline
	\begin{subfigure}{\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../img/plots/cog_hard_baseline_labels.pdf}
	\end{subfigure}%
	\caption{Comparison of test set accuracies of SAMNet (blue) with original results achieved by the baseline model~\cite{yang2018dataset} (dotted gray) on Canonical (top) and Hard (bottom) variants of the COG dataset.}
	\label{fig:samnet_cog_detailed}\vspace{-10pt}
\end{figure}

@@

In order to quantify SAMNet's feature transfer ability, we used the CoGenT variants of CLEVR and performed a set of experiments, starting from training on CoGenT-A, followed by:
an immediate test (zero-shot learning) on CoGenT-B; and fine-tuning for a single epoch on 30k samples from CoGenT-B before testing (following the methodology used in~\cite{johnson2017inferring, mascharka2018transparency, perez2018film, marois2018transfer}).
In \cref{fig:CoGenT-B-results}, we compare the achieved performance with the previously selected state-of-the-art models.
We also consider a direct test on CoGenT-A; and a test on CoGenT-A after fine-tuning on CoGenT-B.

@@

\begin{figure}[b!]
	\centering
	\includegraphics[width=0.8\textwidth]{../img/plots/cog_temporal_transfer_baselines.pdf}
	\caption{Temporal transfer capabilities of SAMNet (blue) and baseline models (gray -striped and -dotted) when moving from Canonical to Hard  in COG.}
	\label{fig:samnet_cog_overall_transfer}
\end{figure}

### 2020-02-14 Transfer_Learning/arxiv_version/submission_folder/v2/experiments_reasoning_transfer.tex
arxiv v2 submission files

We first trained a single SAMNet instance on CoGenT-A and measured its accuracy on each of the task group $t$ separately.
Those results, indicated in \cref{fig:cogent_reasoning_transfer} as "Train on All", serve as baselines for the next two sets of experiments.

We next trained and tested 5 SAMNet instances, each on a single task group $t$ ("Train on t only").
We observe significant accuracy drops (from 18 up to almost 35 points) in all task groups, aside of \textit{Query Attribute}.
We hypothesize \textit{Query Attribute} requires mastering visual grounding and recognizing object attributes, as opposed to the other groups which may not enforce it.
Thus, training on \textit{Query Attribute} helps others, whereas without any training on that particular group, SAMNet battles developing such skill.

Finally, we trained 5 SAMNet instances as follows: for each task group $t$ we trained a single instance on all tasks but $t$, and tested its performance on $t$ only.
As expected, this setup appears challenging, causing severe performance drops, with accuracy on \textit{Count} and \textit{QueryAttribute} dropping to zero.
This stems from the non-overlapping labels spaces of \textit{Count} (digits 0 to 9) and \textit{QueryAttribute} (all attributes names) with the other groups (binary "yes" / "no"). Thus, the model cannot predict these labels in a zero-shot transfer.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}
In order to further investigate whether reasoning transfer is effective in leveraging information gained by training a task family at a higher level of the hierarchy, we also conducted a set of experiments using the Canonical variant of COG.
The order of the performed experiments is analogical to the one in the previous section.
The first column (\cref{fig:cog_reasoning_transfer}, "Train on All") is the baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Please note that these results are weighted averages of the accuracies of our model on the Canonical variant tasks (\cref{fig:samnet_cog_detailed}).

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.8\textwidth]{../img/plots/cog_reasoning_transfer.pdf}
	\caption{Accuracies of SAMNet when testing its reasoning transfer capabilities on the Canonical variant of COG.}
	\label{fig:cog_reasoning_transfer}
\end{figure}

Next, we experimented with the "Train on t only" setup, i.e. training and testing each of 5 SAMNet instances on a single task group $t$ (i.e. single leaf of the proposed hierarchy).
For \textit{Basic}, \textit{Spatial} and \textit{Obj-Attr}, the accuracy is matching the baseline ("Train on All"), suggesting that each group contain sufficient information for developing all necessary skills.
Therefore, models trained on these groups may not benefit from any additional training on other task groups.
However, the accuracy drops for \textit{Compary} and \textit{Cognitive}, suggesting the opposite.

Therefore, we have performed a final set of experiments ("Train on Group A/B"), where we trained two SAMNet instances: a) trained on all tasks from \textit{Group A} and tested on each task from the lowest groups (i.e. \textit{Basic}, \textit{Obj-Attr} and \textit{Compare}) separately; b) trained on all tasks from \textit{Group B} and tested separately on \textit{Spatial} and \textit{Cognitive}.
As expected, training on the other tasks from \textit{Group A} enabled the model to significantly improve on the \textit{Compare} task (from 58.6\% to 92.3\%).
A similar improvement (from 68.8\% to 97.1\%) was observed in the second experiment, when trained on \textit{Group B} and tested on \textit{Cognitive}.
Moreover, the achieved accuracy is in fact higher by 1.4 point than the one achieved when training on all tasks (95.7\%).
This suggests the model may have lacked capacity to master questions from all tasks and that training on a smaller subset allowed better performance on this subset.
Finally, we can interestingly note for \textit{Obj-Attr} and \textit{Spatial} that training on the group A/B does not appear to help. When comparing with "Train on t only", we observe slight accuracy drops (around 0.3-0.5 point).
Understanding this phenomenon warrants further investigations on the dependencies between the reasoning required by the particular questions and the transferred reasoning.

>>>>

We first trained a single SAMNet instance on CoGenT-A and measured its accuracy on each of the task group $t$ separately.
Those results, indicated in \cref{fig:cogent_reasoning_transfer} as ``Train on All'', serve as baselines for the next two sets of experiments.

We next trained and tested 5 SAMNet instances, each on a single task group $t$ (``Train on t only'').
We observe significant accuracy drops (from 18 up to almost 35 points) in all task groups, aside of \textit{Query Attribute}.
We hypothesize \textit{Query Attribute} requires mastering visual grounding and recognizing object attributes, as opposed to the other groups which may not enforce it.
Thus, training on \textit{Query Attribute} helps others, whereas without any training on that particular group, SAMNet battles developing such skill.

Finally, we trained 5 SAMNet instances as follows: for each task group $t$ we trained a single instance on all tasks but $t$, and tested its performance on $t$ only.
As expected, this setup appears challenging, causing severe performance drops, with accuracy on \textit{Count} and \textit{QueryAttribute} dropping to zero.
This stems from the non-overlapping labels spaces of \textit{Count} (digits 0 to 9) and \textit{QueryAttribute} (all attributes names) with the other groups (binary ``yes'' / ``no''). Thus, the model cannot predict these labels in a zero-shot transfer.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}
\begin{figure}[b!]
	\centering
	\includegraphics[width=0.8\textwidth]{../img/plots/cog_reasoning_transfer.pdf}
	\caption{Accuracies of SAMNet when testing its reasoning transfer capabilities on the Canonical variant of COG.}
	\label{fig:cog_reasoning_transfer}
\end{figure}

In order to further investigate whether reasoning transfer is effective in leveraging information gained by training a task family at a higher level of the hierarchy, we also conducted a set of experiments using the Canonical variant of COG.
The order of the performed experiments is analogical to the one in the previous section.
In~\cref{fig:cog_reasoning_transfer}, the first bar in each cluster, ``Train on All'', is the baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Note that these results are weighted averages of the accuracies of our model on the Canonical variant tasks (\cref{fig:samnet_cog_detailed}).

Next, we experimented with the ``Train on t only'' setup, i.e. training and testing each of 5 SAMNet instances on a single task group $t$ (i.e. single leaf of the proposed hierarchy).
For \textit{Basic}, \textit{Spatial} and \textit{Obj-Attr}, the accuracy is matching the baseline (``Train on All''), suggesting that each group contain sufficient information for developing all necessary skills.
Therefore, models trained on these groups may not benefit from any additional training on other task groups.
However, the accuracy drops for \textit{Compary} and \textit{Cognitive}, suggesting the opposite.

Therefore, we have performed a final set of experiments (``Train on Group A/B''), where we trained two SAMNet instances: a) trained on all tasks from \textit{Group A} and tested on each task from the lowest groups (i.e. \textit{Basic}, \textit{Obj-Attr} and \textit{Compare}) separately; b) trained on all tasks from \textit{Group B} and tested separately on \textit{Spatial} and \textit{Cognitive}.
As expected, training on the other tasks from \textit{Group A} enabled the model to significantly improve on the \textit{Compare} task (from 58.6\% to 92.3\%).
A similar improvement (from 68.8\% to 97.1\%) was observed in the second experiment, when trained on \textit{Group B} and tested on \textit{Cognitive}.
Moreover, the achieved accuracy is in fact higher by 1.4 point than the one achieved when training on all tasks (95.7\%).
This suggests the model may have lacked capacity to master questions from all tasks and that training on a smaller subset allowed better performance on this subset.
Finally, we can interestingly note for \textit{Obj-Attr} and \textit{Spatial} that training on the group A/B does not appear to help. When comparing with ``Train on t only'', we observe slight accuracy drops (around 0.3-0.5 point).
Understanding this phenomenon warrants further investigations on the dependencies between the reasoning required by the particular questions and the transferred reasoning.

### 2020-02-14 Transfer_Learning/arxiv_version/submission_folder/v2/summary.tex
arxiv v2 submission files

Our experiments on transfer learning showed the shortcomings of existing approaches, especially for video reasoning.  Hence, we designed a novel Memory-Augmented Neural Network model called SAMNet, with mechanisms to address these deficiencies.
SAMNet showed significant improvements over SOTA models on the COG dataset for Video Reasoning  and achieved comparable performance on the CLEVR dataset for Image Reasoning.
It also demonstrated excellent generalization capabilities for temporal and reasoning transfer. Moreover, through the cautious use of fine-tuning, SAMNet's performance advanced even further.
We hope that the proposed taxonomy, newly established reasoning transfer tasks, along with the provided baselines will bolster new research on both models and datasets for transfer learning on visual reasoning.

>>>>

Our experiments on transfer learning showed the shortcomings of existing approaches, especially for video reasoning.  Hence, we designed a novel Memory-Augmented Neural Network model called SAMNet, with mechanisms to address these deficiencies.
SAMNet showed significant improvements over state-of-the-art models on the COG dataset for Video Reasoning  and achieved comparable performance on the CLEVR dataset for Image Reasoning.
It also demonstrated excellent generalization capabilities for temporal and reasoning transfer. Moreover, through the cautious use of fine-tuning, SAMNet's performance advanced even further.
We hope that the proposed taxonomy, newly established reasoning transfer tasks, along with the provided baselines will bolster new research on both models and datasets for transfer learning on visual reasoning.
